Theories of deliberated choice

### 2024-05-13 simple.tex
Pbl stated
some attempts at counter-examples

\begin{conjecture}
  Consider a mapping ${\allleadstog} \in (P^{\allhists})^I$. 
  Define $E$ as the equivalence classes in $I$ that considers as equivalent all $i$ mapped to the same element of $P^{\allhists}$.
  Recall that the bijections over $I$ are the subset of $I^I$ that are bijections per $e \in E$, thus $s \in I^I$ is a bijection over $I$ iff $\forall e \in E: \restr{s}{e}$ bijects with its image.

  A claim $C$ is not about types iff $\exists {\allleadstog} \in C, s \in I^I \suchthat ({\allleadstog} \circ s) \notin C$.
\end{conjecture}
\begin{proof}
  Consider a claim $C$ about types $S$,
  thus,
  $C = S^I = \set{{\allleadstog} \in (P^{\allhists})^I \suchthat {\allleadstog}(I) \wsub S}$.
  Consider any ${\allleadstog} \in C$ and $s \in I^I$.
  As ${\allleadstog} \in C$, we have ${\allleadstog}(I) \wsub S$.
  Also, ${\allleadstog}(s(I)) \wsub {\allleadstog}(I)$. 
  Thus, ${\allleadstog}(s(I)) \wsub S$.
  Thus, $({\allleadstog} \circ s) \in C$.

  About the converse, given $C$, 
  define $S = \bigcup_{{\allleadstog} \in C}({\allleadstog}(I))$,
  and assuming that 
  $[\forall {\allleadstog} \in C, s: ({\allleadstog} \circ s) \in C]$, let us show that 
  $C = \set{{\allleadstog} \in (P^{\allhists})^I \suchthat {\allleadstog}(I) \wsub S}$.
  First, considering any ${\allleadstog} \in C$, we have ${\allleadstog}(I) \wsub S$ by definition of $S$.
  Second, considering any ${\allleadstog} \in (P^{\allhists})^I \suchthat {\allleadstog}(I) \in S$,

\end{proof}
\begin{conjecture}
  Consider a claim that is not about types but is anonymous. What’s that?
  Example: the claim “everyone has the same behavior”.

  Example 2, with $\card{I} = 3$: “exactly two individuals have behavior 1”.

  Is there an example with infinitely many behaviors?
  The claim must be not about types. If non empty, the claim includes some mapping in $(P^{\allhists})^I$. If not about types, some such mapping has not all correspondances shuffling $I$ in any number but keeping the same range?

  Is it that if there is a claim anonymous, holding, and fals., then there is a claim about types that is anonymous, holds, and fals?
\end{conjecture}
\begin{theorem}
  Consider the attitude above. Is there a claim that holds and is anonymous and falsifiable?
  Check \cref{ex:falsAnonIfBig}.
\end{theorem}

>>>>

\begin{theorem}
  Given a claim $C$, define $S_C \wsub P^{\allhists}$ as $S_C = \bigcup_{{\allleadstog} \in C}({\allleadstog}(I))$.
  A claim $C$ is about types iff $C = (S_C)^I$.
\end{theorem}
\begin{proof}
  If the consequent holds, then $C$ is about types, by definition.
  If a claim is about types $S$, then consider any $f \in S$, we have $C = S^I$ thus for ${\allleadstog}$ defined as ${\allleadstog}(I) = \set{f}$ we have ${\allleadstog} \in C$, thus $f \in S_C$; and consider any $f \in S_C$, consider any ${\allleadstog} \in C \suchthat f \in {\allleadstog}(I)$, we have ${\allleadstog} \in S^I$, thus $f \in S$.
\end{proof}

\begin{theorem}
  Consider a falsifiable and anonymous claim $C$ such that $\exists {\allleadstog} \in C, f_1, f_2 \suchthat {\allleadstog}(I) = \set{f_1, f_2} \land \card{\set{i \in I \suchthat {\allleadstog}(i) = f_1}} \notin \N \land \card{\set{i \in I \suchthat {\allleadstog}(i) = f_2}} \notin \N$.
  Then, $\forall {\allleadstog} \suchthat {\allleadstog}(I) \wsub \set{f_1, f_2}: {\allleadstog} \in C$.
\end{theorem}
\begin{proof}
  Consider any anonymous claim $C$ satisfying the antecedent and any ${\allleadstog'} \suchthat {\allleadstog'}(I) \wsub \set{f_1, f_2} \land {\allleadstog'} \notin C$. 
  To show that $C$ is not falsifiable,
  consider any observable subset of $\allleadstog'$ and let us show that it is completable in $C$.

  From the antecedent we see that $I$ is infinite.
  Define $I_k$ as the individuals that the observable interrogates and satisfying ${\allleadstog'}(i) = f_k$.
  Consider sets of same sizes $I'_1$ and $I'_2$ among the non observed individuals (meaning those in $I \setminus I_1 \setminus I_2$) and with corresponding types in ${\allleadstog}$ (so that $\forall i \in I'_k: {\allleadstog}(i) = f_k$).
  Define a bijection $\sigma$ that sends $I_1$ to $I'_1$ and $I_2$ to $I'_2$, and conversely, and that sends the elements of $I \setminus (\cup_k I_k) \setminus (\cup_k I'_k)$ to themselves.
  The resulting relation ${\allleadstog} \circ \sigma$ is in $C$ and is compatible with the observable.
\end{proof}

\begin{conjecture}
  If there is a claim anonymous, holding, and fals., then there is a claim about types that holds and is falsifiable.
\end{conjecture}
\begin{proof}
  For a counter-example, we need a situation such that the closure for falsifiability of $\set{{\allleadsto} \circ \sigma}$ is not the whole of $\allleadstos$ and $R({\allleadsto}(I)) = P^{\allhists}$.
  Equivalently, we need a situation such that the closure for falsifiability of $\set{{\allleadsto} \circ \sigma}$ is not the whole of $\allleadstos$ and $\forall \hist \in \allhists: P^{\intvl{\hist}} \wsub \restr{{\allleadsto}(I)}{\intvl{\hist}}$.

  A counter-example could be a situation where exactly one individual given $\emptyset$ chooses $\phi$, with the claim: “there is at most one individual who given $\emptyset$ chooses $\phi$”.
  Nope, this seems to admit a claim about types, as then there is finitely many individuals satisfying $\emptyset$ to $\phi$ and hence some of these functions can be excluded.

  The interesting case is a situation including infinitely many behaviors (not necessarily all), otherwise a claim about types exists.
  In such a case, can we find a true claim that is anonymous and fals but not about types?

  Consider a true claim $C$. It contains the true situation, thus involving $f_1, …$. Say it includes infinitely many $f_1$ and exactly one $f_2$.
  A hypothetical situation with infinitely many $f_1$ and exactly two $f_2$ can be distinguished.

  Consider the claim “there is exactly one behavior of each kind”. That is not fals because excludes the possibility that $f_1$ is not there (and if it’s not, impossible to observe). Even if $C$ admits that any finite number of behaviors may miss, it’s still not fals: if all odd $f$ miss (we have all even $f$ and no more), we will never prove it.
  In fact, if an infinite subset of the support is there, it is not empirically distinguishable from all the subset is there; thus the claim must admit for all infinite strict subsets as well.
  Thus, the claim could be “there is at most one behavior of each kind”.
  “Kind” here must be: satisfying a given function in $P^{\intvl{\hist}}$ (more is not counter-observable).

  Thus, consider the claim that $\forall \hist, f \in P^{\intvl{\hist}}: $ there is at one one behavior satisfying $f$. Well, that is self-contradicting.

  The claim must be not about types. If non empty, the claim includes some mapping in $(P^{\allhists})^I$. If not about types, some such mapping has not all correspondances picking from various elements in $C$. But it is also not simply a bijection over $I$: either it picks from multiple elements of $C$, or it picks in different numbers.

  In fact, if an anonymous and falsifiable claim includes a relation that includes an infinite number of $f_1$ and an infinite number of $f_2$ and so on, then I believe that the claim must also include the relations that include any finite number of $f_1$, because that case is impossible to distinguish empirically.

  Check also \cref{ex:falsAnonIfBig}.
\end{proof}
